\documentclass[12pt]{amsart}

\textwidth = 6.2 in
\textheight = 8.5 in
\oddsidemargin = 0.0 in
\evensidemargin = 0.0 in
\topmargin = 0.0 in
\headheight = 0.0 in
\headsep = 0.3 in
\parskip = 0.05 in
\parindent = 0.3 in

\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{color}
\def\cc{\color{blue}}
\usepackage[normalem]{ulem}
\usepackage{amsfonts, amsmath, amssymb, amsthm}
\usepackage{systeme}
\usepackage[none]{hyphenat}
\usepackage{graphicx}
\graphicspath{{./images/}}
\usepackage{esint}
\usepackage{cancel}
\usepackage{physics}

\title{Homework 12}
\author{Sai Sivakumar}

\newtheorem{theorem}            {Theorem}[section]
\newtheorem{proposition}        [theorem]{Proposition}

\newcommand{\RR}{\mathbb{R}}

\begin{document}
\maketitle

Suppose $f:[a,b]\to\RR$ is bounded. Let
\[
 C= \sup\{|f(x)|: x\in [a,b]\}.
\]

\begin{enumerate}[(i)]\itemsep=10pt
 \item Show, for any $a\le x,y\le b,$ that
\[
 |f(x)^2-f(y)^2| \le 2C |f(x)-f(y)|.
\]
\item Fix $a\le c<d \le b$ and let 
 $m,M$ denote the infimum and supremum 
 of the set $\{f(x): x\in [c,d]\}$ respectively.
 Show, for any $c\le x,y\le d,$ that
\[
  |f(x)^2-f(y)^2| \le 2C(M-m).
\]
 \item Let $m^*,M^*$ denote the infimum and supremum
 of the set $\{f(x)^2: x\in [c,d]\}.$ Prove
\[
 M^*-m^* \le 2C (M-m).
\]
 \item Prove if $f$ is integrable, then so is $f^2.$
\end{enumerate}
\bigskip
I spoke with Nicholas Kapsos about details of the proof.
\begin{proof} Let $f\colon [a,b]\to \mathbb{R}$ be bounded with $C = \sup\{|f(x)|: x\in [a,b]\}$ as given. 
\begin{enumerate}[(i)]
    \item For $x,y\in [a,b]$, \begin{align*}
        \abs{f(x)^2 - f(y)^2} &= \abs{f(x)+f(y)}\abs{f(x)-f(y)}\\
        &\leq (\abs{f(x)} + \abs{f(y)})\abs{f(x)-f(y)}\\
        &\leq 2C\abs{f(x)-f(y)}.
    \end{align*}
    \item Fix $a\le c<d \le b$ and let $m$ and $M$ denote the infimum and supremum of the set $\{f(x): x\in [c,d]\}$ respectively. We have $m\leq f(x), f(y) \leq M$, for $x,y\in [c,d]$. It follows that $f(x) - f(y) \leq M - m$, and by taking the absolute value $\abs{f(x) - f(y)}\leq M-m$. Then \[\abs{f(x)^2 - f(y)^2} \leq 2C\abs{f(x)-f(y)} \leq 2C(M-m).\]
    \item Let $m^*$ and $M^*$ denote the infimum and supremum
    of the set $\{f(x)^2: x\in [c,d]\}$ respectively. Then given any $\eta > 0$, there exists $u,v\in [c,d]$ such that \begin{align*}
        M^{\ast} - \eta &\leq f(u)^2\\
        m^{\ast} + \eta &\geq f(v)^2,
    \end{align*} so that $M^{\ast} - m^{\ast} - 2\eta \leq f(u)^2 - f(v)^2$. Since $m^{\ast}\leq M^{\ast}$, it follows that \begin{align*}
        M^{\ast} - m^{\ast} &\leq f(u)^2 - f(v)^2 + 2\eta\\
        &\leq \abs{f(u)^2 - f(v)^2} + 2\eta\\
        &\leq 2C(M-m) + 2\eta.
    \end{align*} Since $\eta>0$ was arbitrary, $M^{\ast} - m^{\ast}\leq 2C(M-m)$.
    \item Suppose $f$ is integrable. Since the space of integrable functions is a vector space, it follows that $2Cf$ is integrable. Then given $\varepsilon>0$, there is a partition $P = \{a = x_0, x_1, \dots, x_{n-1}, x_n = b\}$ such that $U(2Cf, P) - L(2Cf, P)\leq \varepsilon$.
    
    Let $m_i^*$ and $M_i^*$ denote the infimum and supremum
    of the set $\{f(x)^2: x\in [x_i,x_{i+1}]\}$ respectively and let $m_i$ and $M_i$ denote the infimum and supremum of the set $\{f(x): x\in [x_i,x_{i+1}]\}$ respectively. Then by the result of (iii), we have \begin{align*}
        U(f^2, P) - L(f^2,P) &= \sum_i (M^{\ast}_i - m^{\ast}_i)(x_{i+1} - x_i)\\
        &\leq \sum_i 2C(M_i-m_i)(x_{i+1} - x_i)\\
        &= U(2Cf, P) - L(2Cf, P)\leq \varepsilon.
    \end{align*}

    Hence $f^2$ is integrable.
    \end{enumerate}
\end{proof}
\end{document}