\documentclass[12pt]{amsart}

\textwidth = 6.2 in
\textheight = 8.5 in
\oddsidemargin = 0.0 in
\evensidemargin = 0.0 in
\topmargin = 0.0 in
\headheight = 0.0 in
\headsep = 0.3 in
\parskip = 0.05 in
\parindent = 0.3 in

\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{color}
\def\cc{\color{blue}}
\usepackage[normalem]{ulem}
\usepackage{amsfonts, amsmath, amssymb, amsthm}
\usepackage{systeme}
\usepackage[none]{hyphenat}
\usepackage{graphicx}
\graphicspath{{./images/}}
\usepackage{esint}
\usepackage{cancel}
\usepackage{physics}

\title{Homework 2}
\author{Sai Sivakumar}

\newtheorem{theorem}            {Theorem}[section]
\newtheorem{proposition}        [theorem]{Proposition}

\newcommand{\RR}{\mathbb{R}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\DD}{\mathbb{D}}
\newcommand{\cS}{\mathcal{S}}

\begin{document}
\maketitle

\thispagestyle{empty}
The sequence of function $D_N:\RR\to \CC,$  
\[
 D_N(t) = \sum_{n=-N}^N e^{int}
\]
 is the Dirichlet kernel. The sequence of functions $K_N:\RR\to \CC,$ 
\[
 K_N(t) =   \frac{1}{N+1}\sum_{n=0}^N D_n(t) 
\]
 is Fejer's kernel. It satisfies
\begin{enumerate}[(i)]
 \item $K_n(t)\ge 0$ for all $n$ and $t;$
 \item $K_N$ is $2\pi$-periodic;
 \item  $\frac{1}{2\pi}\, \int_{-\pi}^{\pi} K_n(t)\, dt = 1$ for all $n;$ 
 \item  for each $\delta>0,$
\[
  \lim_{n\to\infty} \int_{\delta \le |t|<\pi} K_n(t)\, dt \, = \, 0.
\]
\end{enumerate}
\bigskip

Let $\overline{\DD}=\{|z| \le 1\}.$ Suppose $f:\overline{\DD}\to \CC$ is continuous and $f|_{\DD}:\DD\to\CC$ is analytic. Follow the outline below
(or otherwise) prove $f$ is uniformly approximable in $\DD$  by (analytic)  polynomials.

\begin{enumerate}[(A)]\itemsep=10pt
 \item  Show, with 
 $z=re^{i\theta}$ and $0\le r<1,$ that
\[
 p_n(z) =\frac{1}{2\pi} \int_{-\pi}^\pi f(re^{i(\theta -t)})\, K_n(t)\, dt
  = \frac{1}{2\pi} \int_{-\pi}^\pi f(ze^{-i\, t})\, K_n(t)\, dt
\]
 is a polynomial  (of degree at most $n$). Suggestion: Expand $f$ in a power series.
\item Show $(p_n)$ converges uniformly to $f$ on $\overline{\DD}.$ Suggestion: Note
\[
 f(z) =\frac{1}{2\pi}\int_{-\pi}^\pi f(z) K_n(t)\, dt 
\]
and, since $f$ is uniformly continuous, one can uniformly estimate, in  $\DD,$ 
\[
    |f(z)-f(ze^{-it})|
\]
 by choosing $|t|$ sufficiently small.  
 %to take advantage of properties of the sequence $(K_n)$; 
%and further,
\end{enumerate}

 \bigskip

\begin{proof}
\baselineskip=24pt
For each $n\geq 0$ we have \[K_n(t) = \frac{1}{n+1}\left[e^{-i(n+1)t}+ \cdots+ ne^{-it} + (n+1) + ne^{it} +\cdots + e^{i(n+1)t}\right],\] and for $t\in[-\pi,\pi]$ we have some uniform bound $\abs{K_n(t)}\leq C_n$.

As $f|_\DD$ is analytic it has a power series representation \[f(z) = \sum_{k=0}^\infty a_kz^k\] which converges uniformly in $\DD$. The sequence of partial sums $f_K\colon \mathbb{D}\to\mathbb{C}$ given by \[f_K(z) = \sum_{k=0}^Ka_kz^k\] is uniformly bounded above in $K$ and $z$ due to the uniform convergence of the power series for $f$ in $\DD$. [Uniform convergence is convergence in the $\sup$ norm, and convergent sequences are bounded.] 

For $z = re^{i\theta}$ with $0\leq r < 1$ and for $-\pi\leq t\leq \pi$ we have $ze^{-it}\in \DD$. By the dominated convergence theorem [via the uniform bounds obtained above] we have \[\lim_{K\to\infty}\frac{1}{2\pi}\int_{-\pi}^{\pi}f_K(ze^{-it})K_n(t)\dd{t} = \frac{1}{2\pi}\int_{-\pi}^{\pi}f(ze^{-it})K_n(t)\dd{t}.\] But for $K\geq n$, by integrating each term in \begin{multline*}
\frac{1}{2\pi}\int_{-\pi}^{\pi}f_K(ze^{-it})K_n(t)\dd{t} \\ = \frac{1}{2\pi}\int_{-\pi}^{\pi}\left(\sum_{k=0}^Ka_kz^ke^{-ikt}\right)\left(\frac{1}{n+1}\left[e^{-i(n+1)t}+ \cdots+ ne^{-it} + (n+1) + ne^{it} +\cdots + e^{i(n+1)t}\right]\right)\dd{t}\end{multline*}
we find that the highest degree of $z^k$ that survives is $k=n$. Then in the limit we also have that only up to the $z^n$ term survives. Hence $p_n(z) = \frac{1}{2\pi}\int_{-\pi}^{\pi}f(ze^{-it})K_n(t)\dd{t}$ is a polynomial of degree at most $n$, for each $n\geq 0$.

Let $\varepsilon>0$ be given. Due to the uniform continuity of $f$ in $\overline{\DD}$, there is $\delta>0$ such that for $z\in\overline{\DD}$ and $\abs{t}$ small enough so that $\abs{z-ze^{-it}}<\delta$, $|f(z)-f(ze^{-it})|<\varepsilon/2$ (rough estimates using the power series for sine and cosine should give that $\abs{t}$ could be taken proportional to $\delta$ when $\delta$ is small). Now take $N$ large enough so that for $n\geq N$ we have $M\int_{\delta \leq |t|<\pi} K_n(t)\dd{t}<\varepsilon/2$, where $M$ is a uniform upper bound of $|f(z)-f(ze^{-it})|$ for all $-\pi\leq t\leq \pi$. [So $N$ only depends on $\delta$, which only depends on $\varepsilon$.] Then for $n\geq N$, we have \begin{align*}
  \abs{ f(z)-p_n(z)} &= \abs{\frac{1}{2\pi}\int_{-\pi}^{\pi}(f(z)-f(ze^{-it}))K_n(t)\dd{t}}\\
  &\leq \frac{1}{2\pi}\int_{-\pi}^{\pi}\abs{f(z)-f(ze^{-it})}K_n(t)\dd{t}\\
  &\leq \frac{\varepsilon}{4\pi}\int_{|t|<\delta} K_n(t)\dd{t} + \frac{M}{2\pi}\int_{\delta \leq |t|<\pi} K_n(t)\dd{t}\\
  &\leq \frac{\varepsilon}{2\pi} .
\end{align*} Hence $f$ may be uniformly approximated by polynomials in $\DD$.
\end{proof}
\end{document}