\documentclass[11pt]{article}

% packages
\usepackage{physics}
% margin spacing
\usepackage[top=1in, bottom=1in, left=0.5in, right=0.5in]{geometry}
\usepackage{hanging}
\usepackage{amsfonts, amsmath, amssymb, amsthm}
\usepackage{systeme}
\usepackage[none]{hyphenat}
\usepackage{fancyhdr}
\usepackage[nottoc, notlot, notlof]{tocbibind}
\usepackage{graphicx}
\graphicspath{{./images/}}
\usepackage{float}
\usepackage{siunitx}
\usepackage{esint}
\usepackage{cancel}
\usepackage{enumitem}
\usepackage{quiver}

% permutations (second line is for spacing)
\usepackage{permute}
\renewcommand*\pmtseparator{\,}

% colors
\usepackage{xcolor}
\definecolor{p}{HTML}{FFDDDD}
\definecolor{g}{HTML}{D9FFDF}
\definecolor{y}{HTML}{FFFFCF}
\definecolor{b}{HTML}{D9FFFF}
\definecolor{o}{HTML}{FADECB}
%\definecolor{}{HTML}{}

% \highlight[<color>]{<stuff>}
\newcommand{\highlight}[2][p]{\mathchoice%
  {\colorbox{#1}{$\displaystyle#2$}}%
  {\colorbox{#1}{$\textstyle#2$}}%
  {\colorbox{#1}{$\scriptstyle#2$}}%
  {\colorbox{#1}{$\scriptscriptstyle#2$}}}%

% header/footer formatting
\pagestyle{fancy}
\fancyhead{}
\fancyfoot{}
\fancyhead[L]{MAS6332 Algebra}
\fancyhead[C]{Homework 5}
\fancyhead[R]{Sai Sivakumar}
\fancyfoot[R]{\thepage}
\renewcommand{\headrulewidth}{1pt}

% paragraph indentation/spacing
\setlength{\parindent}{0cm}
\setlength{\parskip}{10pt}
\renewcommand{\baselinestretch}{1.25}

% extra commands defined here
\newcommand{\br}[1]{\left(#1\right)}
\newcommand{\sbr}[1]{\left[#1\right]}
\newcommand{\cbr}[1]{\left\{#1\right\}}

\newcommand{\dprime}{\prime\prime}

% bracket notation for inner product
\usepackage{mathtools}

\DeclarePairedDelimiterX{\abr}[1]{\langle}{\rangle}{#1}

\DeclareMathOperator{\Span}{span}
\DeclareMathOperator{\nullity}{nullity}
\DeclareMathOperator\Aut{Aut}
\DeclareMathOperator\Inn{Inn}
\DeclareMathOperator{\Orb}{Orb}
\DeclareMathOperator{\lcm}{lcm}
\DeclareMathOperator{\Hol}{Hol}
\DeclareMathOperator{\Jac}{Jac}
\DeclareMathOperator{\rad}{rad}
\DeclareMathOperator{\Tor}{Tor}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\Gal}{Gal}
\DeclareMathOperator{\Nat}{Nat}
\DeclareMathOperator{\Frac}{Frac}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Ext}{Ext}
\DeclareMathOperator{\aug}{aug}

% set page count index to begin from 1
\setcounter{page}{1}

\begin{document}
\subsection*{Graded}
\begin{enumerate}
    \item ($G$-modules with Trivial Cohomology) Let $G$ be a finite group and $A$ a $G$-module. Show that there exists a $G$-module $B$ with an injective map of $G$-modules $A\hookrightarrow B$ such that $H^n(G,B) = 0$ for $n>0$. (We say $B$ is cohomologically trivial for $G$.) Here is a suggested strategy, which is a streamlined version of the arguments in Section 17.2. \begin{itemize}
        \item Take $B = \Hom_{\mathbb{Z}}(\mathbb{Z}G, A)$, with $G$ acting on the codomain.
        \item Use tensor-hom adjunction to show that for any $\mathbb{Z}G$-module $P$, $\Hom_{\mathbb{Z}G}(P,B)\cong \Hom_\mathbb{Z}(P,A)$.
        \item Show $\Ext_{\mathbb{Z}G}^n(\mathbb{Z},B)\cong \Ext_\mathbb{Z}^n(\mathbb{Z},A)$ by applying $\Hom_{\mathbb{Z}G}(-,B)$ to a projective resolution of $\mathbb{Z}G$-modules for $\mathbb{Z}$.
    \end{itemize} \begin{proof}
        With $G$ a finite group and $A$ a $G$-module as above let $B = \Hom_{\mathbb{Z}}(\mathbb{Z}G,A)$, where the $G$-action is pointwise (so $gf$ is given by $(gf)(x) = gf(x)$). By the tensor-hom adjunction we have for any $\mathbb{Z}G$-module $P$ that \[\Hom_{\mathbb{Z}G}(P,B) = \Hom_{\mathbb{Z}G}(P,\Hom_{\mathbb{Z}}(\mathbb{Z}G,A))\cong \Hom_{\mathbb{Z}}(P\otimes_{\mathbb{Z}G}\mathbb{Z}G, A) \cong \Hom_\mathbb{Z}(P,A).\]

        Now let $\cbr{P_i}$ be a projective resolution of $\mathbb{Z}G$-modules for $\mathbb{Z}$: \[\cdots\to P_n\to\cdots\to P_1\to P_0\to \mathbb{Z}\to 0\] Apply the $\Hom_{\mathbb{Z}G}(-,B)$ functor to obtain the cochain complex (removing the first group) \[0\to \Hom_{\mathbb{Z}G}(P_0,B)\xrightarrow{d_1} \Hom_{\mathbb{Z}G}(P_1,B)\xrightarrow{d_2}\cdots\xrightarrow{d_n} \Hom_{\mathbb{Z}G}(P_n,B)\xrightarrow{d_{n+1}}\cdots.\] The cohomology of this complex is given by the groups $\Ext_{\mathbb{Z}G}^i(\mathbb{Z},B) = \ker{d_{i+1}}/\im{d_i}$. Using the tensor-hom adjunction we produce isomorphisms $f_i\colon \Hom_{\mathbb{Z}G}(P_i,B)\to \Hom_\mathbb{Z}(P_i,A)$; produce the cochain complex (is still a cochain complex since the $f_i$ are isomorphisms) \[0\to \Hom_\mathbb{Z}(P_0,A)\xrightarrow{f_1d_1f_0^{-1}} \Hom_\mathbb{Z}(P_1,A)\xrightarrow{f_2d_2f_1^{-1}}\cdots\xrightarrow{f_nd_nf_{n-1}^{-1}} \Hom_\mathbb{Z}(P_n,A)\xrightarrow{f_{n+1}d_{n+1}f_n^{-1}}\cdots.\] The cohomology of this complex is isomorphic to the cohomology of the previous complex given since the $f_i$ are isomorphisms: we have $\ker{d_{i+1}}\cong \ker(f_{i+1}d_{i+1}f_i^{-1})$ and $\im{d_i}\cong \im(f_id_if_{i-1}^{-1})$. So $\Ext_{\mathbb{Z}G}^n(\mathbb{Z},B)\cong \Ext_\mathbb{Z}^n(\mathbb{Z},A)$ as needed. But for $n>0$, $\Ext_\mathbb{Z}^n(\mathbb{Z},A)$ since $\mathbb{Z}$ is a projective (free) $\mathbb{Z}$-module.
    \end{proof}
    \item (17.4.10) Suppose $\mathbb{F}_q$ is a finite field with $G= \Gal(\mathbb{F}_{q^d}/\mathbb{F}_q) = \abr{\sigma_q}$ where $\sigma_q$ is the Frobenius automorphism, and let $N$ be the usual norm element for the cyclic group $G$. \begin{enumerate}
        \item Use Hilbert's Theorem 90 to prove that $|{}_N(\mathbb{F}_{q^d}^\times)| = (q^d-1)/(q-1)$, and deduce that the norm map from $\mathbb{F}_{q^d}$ to $\mathbb{F}_q$ is surjective. \begin{proof}
        Let $F = \mathbb{F}_q$ and $K = \mathbb{F}_{q^d}$, so that $G = \Gal(K/F) = \abr{\sigma_q}$. Hilbert's Theorem 90 states that $H^1(G,K^\times) = {}_NK^\times/(\sigma_q/\id_K)K^\times = 0$ so that $\abs{{}_NK^\times} = \abs{(\sigma_q/\id_K)K^\times}$. For $r\in K^\times$, $r^q/r = 1$ if and only if $r^q = r$; that is, if $r$ is in $(K^G)^\times = \mathbb{F}_q^\times$. It follows that $\abs{(\sigma_q/\id_K)K^\times} = |\mathbb{F}_{q^d}^\times/\mathbb{F}_q^\times| = (q^d-1)/(q-1)$. 

        Then by interpreting ${}_NK^\times$ as the kernel of the norm map of the cyclic groups $K^\times\to F^\times$ we have that the image of the norm map is a (cyclic) subgroup of order $q-1$ in $\mathbb{F}_q^\times$. But $\mathbb{F}_q^\times$ has order $q-1$ so the image of the norm map on all of $\mathbb{F}_{q^d}$ is exactly $\mathbb{F}_q$.
        \end{proof}
        \item Prove that $H^n(G,\mathbb{F}_{q^d}^\times) = 0$ for all $n\geq 1$. \begin{proof}
            We saw earlier that the cohomology of a finite cyclic group alternates between two groups. In our setting we have (with $(K^\times)^G = (K^G)^\times$) \[H^n(G,K^\times) = \begin{cases}
            (K^G)^\times/NK^\times & n\geq 2, \text{ even} \\
            {}_NK^\times/(\sigma_q/\id_K)K^\times & n\geq 1,\text{ odd}\end{cases}.\] By Hilbert's Theorem 90 all the odd index cohomology groups are trivial. For the nonzero even index groups, since the norm map is surjective we have that $NK^\times = \mathbb{F}_q^\times = (K^G)^\times$ from which it follows the nonzero even index cohomology groups are trivial.
        \end{proof}
    \end{enumerate}
\end{enumerate}
\subsection*{Additional Problems}
\begin{enumerate}
    % 17.3.11 or Problem 2
    %\item (17.3.4) Let $A$ be the Klein $4$-group and let $G = \Aut(A)\cong S_3$ act on $A$ in the natural fashion. Prove that $H^1(G,A) = 0$. [Show that in the semidirect product $E = A\rtimes G$, $G$ is the normalizer of a Sylow $3$-subgroup of $E$. Apply Sylow's theorem to show that all complements to $A$ in $E$ are conjugate.]\begin{proof}
    %    In the semidirect product $E = A\rtimes G$ of order $24$, a Sylow $3$-group is given by the cyclic subgroup $\abr{(1,(1\,2\,3))}$. Furthermore, this subgroup 
        %(and any other Sylow $3$-subgroup, as they are conjugate to each other) 
    %    is normalized by $G\cong \abr{(1,(1\,2)),(1,(1\,2\,3))}$ since $\abr{(1\,2\,3)}$ is an index $2$ subgroup of $S_3$. Then Sylow's theorem says that the number of $3$-Sylows is the index of $G$ in $E$, which is $4$.
    %\end{proof}
    \item (17.4.2) Let $A=\mathbb{Z}/4\mathbb{Z}$ and let $G$ be the cyclic group of order $2$ acting trivially on $A$.\begin{enumerate}
        \item Prove that $\abs{C^2(G,A)} = 2^8$. 
        \item Use the coboundary condition to show that $\abs{B^2(G,A)} = 2^3$.
        \item Use the cocycle condition to show that $\abs{Z^2(G,A)} \leq 2^4$. 
        \item Show that $\abs{H^2(G,A)} = 2$ by exhibiting two inequivalent extensions of $G$ by $A$ and their corresponding cocycles.
    \end{enumerate} \begin{proof}
        The number of functions from $G\times G\to A$ is given by $\abs{A}^{\abs{G\times G}} = 4^4 = 2^8$ so $\abs{C^2(G,A)} = 2^8$. The function $f\colon G\times G\to A$ is a coboundary if there is $f_1\colon G\to A$ with $f(g,h) = f_1(h)-f_1(gh)+f_1(g)$. By taking various choices $g,h\in Z_2 = \abr{x}$ we deduce that $f(1,1) = f(1,x) = f(x,1)$ and that $f(x,x) = 2f_1(x)-f(1,1)$. There are four choices for $f(1,1)$ and two choices for $f(x,x)$ since $2f_1(x)$ is either $0$ or $2$. Hence $\abs{B^2(G,A)} = 8 = 2^3$. For $f$ to be a cocycle we must have for all $g,h,k$ that $f(g,h) = f(h,k)+f(gh,k)-f(gh,k)$. By varying $g,h$ we find that $f(1,1) = f(1,k) = f(x,1)$, that $f(1,x) = f(1,xk)$, and finally that $f(x,x) = f(x,k)+ f(x,xk)-f(1,k)$. By varying $k$ obtain the relations $f(x,1) = f(1,1) = f(1,x)$ so that only $f(x,x)$ is undetermined. There are four choices for each of the two so there are at most $16 = 2^4$ elements in $Z^2(G,A)$. There are indeed two inequivalent extensions of $G$ by $A$, given by the direct product $Z_4\times Z_2$ and the cyclic groups $Z_8$ (both have cyclic normal subgroups of order $4$ with quotient $Z_2$). The cocycle for the direct product is the zero cocycle. For the cyclic group we find that $f(x,x) = 2$ and $f(1,x) = f(x,1) = f(1,1)= 0$. 
    \end{proof}
    \item (Artin-Schreier Theory). Let $K/F$ be a Galois extension of fields. As on the final from last semester, linear independence of characters shows that $H^1(\Gal(K/F),K) = 0$ (here $K$ is viewed as a group using addition with the natural action of the Galois group). 
    
    Let $p$ be a prime, $F$ a field of characteristic $p$. Let $K$ be the separable closure of $F$. Using the short exact sequence \[0\to \mathbb{F}_p\to K^\times\xrightarrow{f}K^\times\to 1\] where $f(\alpha) = \alpha^p-\alpha$, adapt the argument for Kummer theory to show that cyclic $\mathbb{Z}/p\mathbb{Z}$-extensions of a field $F$ of characteristic $p$ are in bijection with the non-trivial subgroups of $F^\times/f(F^\times)$. (Compare with the final exam from last semester.) \begin{proof}
        We check that the sequence given above is a short exact sequence: Certainly $\mathbb{F}_p$ is included in $K$ and $f$ is surjective since $x^p-x$ is separable. Furthermore, we know that elements of $\mathbb{F}_p$ satisfy $x^p-x = 0$ and also with $\mathbb{F}_p$ the splitting field for the same polynomial, it follows that we have exactness in the center. 

        Then in the long exact sequence of Galois cohomology, we have $0\to \mathbb{F}_p^G\to (K^\times)^G\xrightarrow{f}(K^\times)^G\to H^1(G,\mathbb{F}_p)\to H^1(G, K^\times) \to \cdots$. Like before we should have that $H^1(G, K^\times)$ is trivial, and with $F$ the fixed field of $K$ (and $\mathbb{F}_p\subset F$ so we also fix $\mathbb{F}_p$) the sequence becomes $0\to \mathbb{F}_p\to F^\times\xrightarrow{f}F^\times\to H^1(G,\mathbb{F}_p)\to 0$. Then $H^1(G,\mathbb{F}_p)$ is isomorphic to the cokernel $F^\times/f(F^\times)$. 

        So $H^1(G,\mathbb{F}_p)$ should be given by the group of continuous homomorphisms from $G$ into $\mathbb{F}_p = \mathbb{Z}/p\mathbb{Z}$. Kernels of nonzero maps in this group are closed normal subgroups, which by Galois theory correspond to cyclic $\mathbb{Z}/p\mathbb{Z}$ extensions as the images of these maps are nonzero (the whole group).
    \end{proof}
\end{enumerate}
\subsection*{Feedback}
\begin{enumerate}
    \item None.
    \item Things seem to be the same I think.
\end{enumerate}
\end{document}